---
title: "Feature Extraction from HR Dataset"
author: "Data Scientist"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

load("/Users/roberto/Desktop/sds_r/data_part2.RData")
names(hw)
# Count how many observation in each zone
table(hw$y)
```

# **Introduction**
This RMarkdown document extracts relevant statistical and time-series features from a dataset containing heart rate (HR) zones, speed, and altitude measurements over a 60-second interval. The processed dataset will be used for classification tasks.

## **1. Load Necessary Libraries**
We use several R packages for data manipulation and statistical feature extraction.

```{r load-libraries}
library(dplyr)      # For data wrangling
library(moments)    # For skewness and kurtosis
library(ggplot2)    # For visualization
```

## **2. Load Dataset**
Assuming the dataset `hw` is already loaded, otherwise, load it from an external file.

```{r load-data}
# load("data_part2.RData")
head(hw) # Display first few rows
```

## **3. Define Feature Extraction Function**
The following function extracts **basic statistical features** from both speed and altitude time series data. These features include central tendency, dispersion, shape, and trend.

```{r feature-extraction}
extract_features <- function(data) {
  data %>%
    rowwise() %>%
    mutate(
      # Basic Speed Statistics
      speed_mean = mean(c_across(starts_with("sp.")), na.rm = TRUE), # Average speed
      speed_sd = sd(c_across(starts_with("sp.")), na.rm = TRUE),     # Speed variability
      speed_min = min(c_across(starts_with("sp.")), na.rm = TRUE),   # Minimum speed recorded
      speed_max = max(c_across(starts_with("sp.")), na.rm = TRUE),   # Maximum speed recorded
      speed_range = speed_max - speed_min,                            # Speed fluctuation range
      
      # Speed Distribution Shape
      speed_skew = skewness(c_across(starts_with("sp.")), na.rm = TRUE),      # Asymmetry in speed data
      speed_kurtosis = kurtosis(c_across(starts_with("sp.")), na.rm = TRUE),  # Peakiness of the speed distribution
      
      # Speed Trend
      speed_slope = ifelse(length(unique(c_across(starts_with("sp.")))) > 1, 
                           coef(lm(c_across(starts_with("sp.")) ~ seq(1, 60)))[2], NA), # Trend over time
      
      # Basic Altitude Statistics
      alt_mean = mean(c_across(starts_with("al.")), na.rm = TRUE), # Average altitude
      alt_sd = sd(c_across(starts_with("al.")), na.rm = TRUE),     # Altitude variability
      alt_min = min(c_across(starts_with("al.")), na.rm = TRUE),   # Minimum altitude recorded
      alt_max = max(c_across(starts_with("al.")), na.rm = TRUE),   # Maximum altitude recorded
      alt_range = alt_max - alt_min,                                # Altitude fluctuation range
      
      # Altitude Distribution Shape
      alt_skew = skewness(c_across(starts_with("al.")), na.rm = TRUE),      # Asymmetry in altitude data
      alt_kurtosis = kurtosis(c_across(starts_with("al.")), na.rm = TRUE),  # Peakiness of altitude distribution
      
      # Altitude Trend
      alt_slope = ifelse(length(unique(c_across(starts_with("al.")))) > 1, 
                         coef(lm(c_across(starts_with("al.")) ~ seq(1, 60)))[2], NA) # Trend over time
    ) %>%
    ungroup()
}
```

## **4. Apply Feature Extraction**
Applying the function to the dataset to create a new processed dataset.

```{r apply-features}
dataset_features <- extract_features(hw)

# Ensure correct selection of columns
dataset_features <- dataset_features %>% select(all_of(c("y", "speed_mean", "speed_sd", "speed_min", "speed_max", "speed_range", 
                                                         "speed_skew", "speed_kurtosis", "speed_slope", "alt_mean", "alt_sd", 
                                                         "alt_min", "alt_max", "alt_range", "alt_skew", "alt_kurtosis", "alt_slope")))
```

## **5. Save Processed Dataset**
```{r save-data}
write.csv(dataset_features, "processed_hw_data.csv", row.names = FALSE)
```

## **6. Exploratory Data Analysis**
### **Summary Statistics**
```{r summary}
summary(dataset_features)
```

### **Visualization of Feature Distributions**
#### **Speed Mean Distribution**
```{r speed-mean-plot}
ggplot(dataset_features, aes(x = speed_mean, fill = as.factor(y))) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of Speed Mean by HR Zone", x = "Speed Mean", y = "Density")
```

#### **Altitude Mean Distribution**
```{r alt-mean-plot}
ggplot(dataset_features, aes(x = alt_mean, fill = as.factor(y))) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of Altitude Mean by HR Zone", x = "Altitude Mean", y = "Density")
```

#### different approach :: 
```{r different-approach}
# Load the data
#load("hw_data.RData")
# Variable names
names(hw)
# Count how many observation in each zone
table(hw$y)

extract_features_one <- function(row_data, speed_cols, altitude_cols) {
  # Estraiamo i valori delle serie "speed" e "altitude"
  sp_values <- as.numeric(row_data[speed_cols])
  al_values <- as.numeric(row_data[altitude_cols])
  
  # (1) Feature su speed
  mean_sp   <- mean(sp_values, na.rm=TRUE)
  median_sp <- median(sp_values, na.rm=TRUE)
  sd_sp     <- sd(sp_values, na.rm=TRUE)
  min_sp    <- min(sp_values, na.rm=TRUE)
  max_sp    <- max(sp_values, na.rm=TRUE)
  range_sp  <- max_sp - min_sp
  diff_sp   <- sp_values[length(sp_values)] - sp_values[1]
  
  # pendenza retta: lm(sp_values ~ tempo)
  time_idx   <- seq_along(sp_values)
  lm_sp      <- lm(sp_values ~ time_idx)
  slope_sp   <- coef(lm_sp)[2]  # coeff angolare
  
  # (2) Feature su altitude
  mean_al   <- mean(al_values, na.rm=TRUE)
  median_al <- median(al_values, na.rm=TRUE)
  sd_al     <- sd(al_values, na.rm=TRUE)
  min_al    <- min(al_values, na.rm=TRUE)
  max_al    <- max(al_values, na.rm=TRUE)
  range_al  <- max_al - min_al
  diff_al   <- al_values[length(al_values)] - al_values[1]
  
  time_idx2  <- seq_along(al_values)
  lm_al      <- lm(al_values ~ time_idx2)
  slope_al   <- coef(lm_al)[2]
  
  # (3) Correlazione speed-altitude (opzionale)
  corr_sp_al <- cor(sp_values, al_values, use="complete.obs")
  
  # Creiamo un vettore con tutte le feature
  feats <- c(
    mean_sp = mean_sp,
    median_sp = median_sp,
    sd_sp = sd_sp,
    min_sp = min_sp,
    max_sp = max_sp,
    range_sp = range_sp,
    diff_sp = diff_sp,
    slope_sp = slope_sp,
    
    mean_al = mean_al,
    median_al = median_al,
    sd_al = sd_al,
    min_al = min_al,
    max_al = max_al,
    range_al = range_al,
    diff_al = diff_al,
    slope_al = slope_al,
    
    corr_sp_al = corr_sp_al
  )
  
  return(feats)
}

extract_features_all <- function(hw) {
  # Troviamo gli indici delle colonne speed/altitude
  speed_cols <- grep("^sp\\.", colnames(hw))
  alt_cols   <- grep("^al\\.", colnames(hw))
  
  # Creiamo un dataframe vuoto per raccogliere le feature
  n <- nrow(hw)
  feature_list <- vector("list", n)
  
  # Cicliamo su ogni riga
  for (i in 1:n) {
    row_data <- hw[i, ]
    feats_i  <- extract_features_one(row_data, speed_cols, alt_cols)
    feature_list[[i]] <- feats_i
  }
  
  # Convertiamo la lista di vettori in un dataframe
  feature_df <- as.data.frame(do.call(rbind, feature_list))
  
  # Aggiungiamo la colonna target y
  feature_df$y <- hw$y
  
  return(feature_df)
}

# ESEMPIO DI USO
feature_data <- extract_features_all(hw)
head(feature_data)

```
## **7. Conclusion**
This document successfully extracts and visualizes relevant features for HR zone classification. The next steps involve applying a classification model and performing hypothesis testing to distinguish between HR zones.